% ============================================================================
% Online Supplementary Information (Online Resource 1)
% Journal: Journal of Regulatory Economics (Springer)
% ============================================================================
%
% Uses the same sn-jnl class as the main paper for consistent Springer styling.
% No line numbers (not needed for supplementary material).
%
\documentclass[pdflatex,sn-basic]{sn-jnl}

% ============================================================================
% PACKAGES
% ============================================================================

\usepackage{amsmath,amssymb}
\usepackage{bbm}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{float}
\usepackage{threeparttable}
\usepackage{array}
\usepackage{longtable}
\usepackage{adjustbox}
\usepackage{pdflscape}
\usepackage{setspace}
\usepackage{hyperref}
\let\orcidlogo\relax  % sn-jnl.cls defines \orcidlogo; clear before orcidlink redefines it
\usepackage{orcidlink}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}

% Figure search path
\graphicspath{{figures/}}

% sn-jnl wraps every \begin{table} in its own threeparttable+center.
% Restore the plain table environment using the aliases the class defines.
\AtBeginDocument{%
  \let\table\tableorg
  \let\endtable\endtableorg
  \let\sidewaystable\sidewaystableorg
  \let\endsidewaystable\endsidewaystableorg
}

% Auto-generated macros (numbers from regression outputs)
\input{paper_macros}

% Appendix table numbering: A1--A4, B1--B4
% \thesection = A, B, ... after \appendix; \thetable resets at each \section.
\renewcommand{\thetable}{\thesection\arabic{table}}
% Fix duplicate hyperref PDF anchors when table counter is reset in section B
\renewcommand{\theHtable}{\thesection\arabic{table}}

% ============================================================================
% DOCUMENT
% ============================================================================

\begin{document}

% Running head for supplementary
\title[Online Supplementary Information]{Online Supplementary Information (Online Resource~1)\\[4pt]
\textit{Broadband Price Elasticity in EU and Eastern Partnership Countries:
Evidence of a Decade-Long Transformation (2010--2024)}}

% ============================================================================
% AUTHORS AND AFFILIATIONS (required by Springer SI guidelines)
% ============================================================================

\author*[1,2,3,4]{\fnm{Samir} \sur{Orujov}\orcidlink{0009-0004-9708-2109}}\email{sorujov@ada.edu.az}

\author[2,5]{\fnm{Ilgar} \sur{Ismayilov}}

\author[2]{\fnm{Jeyhun} \sur{Huseynzade}}

\affil*[1]{\orgname{School of Business, ADA University},
           \orgaddress{\city{Baku}, \country{Azerbaijan}}}

\affil[2]{\orgname{Information Communication Technologies Agency},
           \orgaddress{\city{Baku}, \country{Azerbaijan}}}

\affil[3]{\orgname{CERGE-EI, a joint workplace of Charles University and the Economics Institute of the Czech Academy of Sciences},
           \orgaddress{\city{Prague}, \country{Czech Republic}}}

\affil[4]{\orgname{Universit\'{e} de Bretagne Sud},
           \orgaddress{\city{Vannes}, \country{France}}}

\affil[5]{\orgname{Faculty of Social Sciences, Charles University},
           \orgaddress{\city{Prague}, \country{Czech Republic}}}

% ============================================================================
% ABSTRACT (describes supplementary content)
% ============================================================================

\abstract{This file is the Online Supplementary Information submitted alongside
the article \textit{Broadband Price Elasticity in EU and Eastern Partnership
Countries: Evidence of a Decade-Long Transformation (2010--2024)},
\textit{Journal of Regulatory Economics}.
It contains eight regression tables referenced in the main text as
Online Resource Tables~A1--A4 and B1--B4.
Appendix~A presents the core results: baseline pre-COVID estimates (Table~A1),
COVID-19 interaction effects (Table~A2), robustness to alternative price measures
(Table~A3), and the placebo pre-trend test (Table~A4).
Appendix~B presents additional robustness checks: acute vs.\ post-acute
COVID period decomposition (Table~B1), 2SLS instrumental-variable estimates
(Table~B2), EaP leave-one-country-out jackknife (Table~B3), and alternative
sample restrictions (Table~B4).
All models use two-way fixed effects (country and year) with Driscoll--Kraay
standard errors robust to heteroskedasticity, serial correlation, and
cross-sectional dependence (bandwidth $= 3$).
Significance codes: $^{*}p{<}0.10$, $^{**}p{<}0.05$, $^{***}p{<}0.01$.}

\maketitle

% ============================================================================
% APPENDIX A: BASELINE REGRESSION TABLES
% ============================================================================

\appendix

\section{Baseline Regression Tables}
\label{app:A}
\setcounter{table}{0}

\subsection{Baseline Pre-COVID Results (Table~A1)}

Table~A1 presents baseline two-way fixed effects estimates for the pre-COVID period
(2010--2019) across seven control specifications, from minimal controls (GDP per capita
only, column~1) to the full specification (column~7).

\input{tables/table1_baseline}
\clearpage

\subsection{COVID-19 Interaction Effects (Table~A2)}

Table~A2 presents results from equation~(2) of the main paper, interacting prices with
the COVID-19 period dummy (2020--2024). Column~(1) uses the full 2010--2024 sample with
COVID interactions; columns~(2)--(3) present separate pre-COVID and COVID subsample
estimates for comparison.

\input{tables/table2_covid}
\clearpage

\subsection{Robustness to Alternative Price Measures (Table~A3)}

Table~A3 compares elasticity estimates across three price definitions: income-relative
(GNI\%, baseline), PPP-adjusted, and nominal USD prices. Each panel rows all seven
control specifications for one price measure.

\input{tables/table3_price_robustness}
\clearpage

\subsection{Placebo Test for Pre-COVID Trends (Table~A4)}

Table~A4 presents the placebo test splitting the pre-COVID period into an early
subperiod (2010--2014) and a late subperiod (2015--2019). A significant positive
interaction on the 2015--2019 dummy indicates elasticity was already declining before
the pandemic.

\input{tables/table4_placebo}
\clearpage

% ============================================================================
% APPENDIX B: ADDITIONAL ROBUSTNESS CHECKS
% ============================================================================

\section{Additional Robustness Checks}
\label{app:B}
\setcounter{table}{0}

\subsection{Period-Split: Acute vs.\ Post-Acute COVID (Table~B1)}

Table~B1 decomposes the COVID era into an acute phase (2020--2021) and a post-acute
phase (2022--2024) using a single full-panel interaction model with
price$\,\times\,$period dummies, preserving degrees of freedom. Both sub-periods
exhibit near-zero elasticity for EU and EaP. Panel~B excludes Ukraine and
Belarus---whose data reliability is compromised by war and sanctions---with unchanged
conclusions.

\input{tables/table5_period_split}
\clearpage

\subsection{IV Robustness: 2SLS Estimates (Table~B2)}

Table~B2 presents 2SLS estimates instrumenting fixed broadband price with its own
one-year lag ($\log$ fixed broadband price in GNI\%, lagged one year).

\textbf{Instrument rationale.}
National fixed broadband price baskets are set through annual regulatory and
procurement cycles, generating high year-to-year autocorrelation. The lagged price
strongly predicts the current price (within-demeaned first-stage
$F = \IVFirstF$ pre-COVID; $F = \IVFullF$ full-sample), well above the
Stock--Yogo critical value of 16.38 for 10\% worst-case Wald test size distortion \citep{stock2002testing}.
The exclusion restriction is plausible: conditional on country and year fixed effects
plus time-varying controls, past prices affect current subscriptions only through
current prices, since broadband contracts are monthly or annual and consumers cannot
anticipate next year's regulated tariff when making adoption decisions
today.\footnote{We also tested international internet bandwidth as an alternative
instrument. An over-identified specification combining both instruments strongly
rejected the Hansen $J$ overidentification test ($\chi^2(2) = 262.89$,
$p < 0.001$), indicating bandwidth correlates with connection speed and quality,
which directly enters the demand equation. ITU bandwidth data are also missing for
most EU countries after 2018. Lagged price is therefore the sole preferred
instrument.}

\input{tables/table6_iv_robustness}

\textbf{Results.}
Panel~A shows pre-COVID 2SLS alongside the OLS baseline:
EU$=\IVPreEUB$ (SE$=\IVPreEUSE$, $p=\IVPreEUPval$\IVPreEUStars) and
EaP$=\IVPreEaPB$ (SE$=\IVPreEaPSE$, $p=$\IVPreEaPPval\IVPreEaPStars),
broadly consistent with OLS (EU$=\EUBaseB$, EaP$=\EaPBaseB$), confirming
simultaneity bias is negligible.
Panel~B replicates the full-sample COVID specification:
EU$=\IVFullEUB$ (SE$=\IVFullEUSE$, $p=$\IVFullEUPval\IVFullEUStars) and
EaP$=\IVFullEaPB$ (SE$=\IVFullEaPSE$, $p=$\IVFullEaPPval\IVFullEaPStars),
consistent with the OLS COVID-era attenuation.

\clearpage

\subsection{EaP Jackknife: Leave-One-Country-Out (Table~B3)}

Table~B3 reports leave-one-out estimates dropping each of the six EaP countries in
turn. EaP elasticity ranges from $\EaPJKMax$ to $\EaPJKMin$ across the six
sub-samples. \EaPJKMaxCountry{} contributes the highest price-sensitivity; excluding
it raises the EaP estimate from $\EaPBaseBr$ to $\EaPJKMax$. All six leave-one-out
estimates are significant at the 1\% level, confirming the EaP result is not driven
by any single country.

\input{tables/table7_jackknife}
\clearpage

\subsection{Sample Restrictions (Table~B4)}

Table~B4 reports robustness to alternative sample definitions. The balanced-panel
restriction (countries with complete 2010--2019 data) yields identical estimates,
confirming attrition does not drive the results. Excluding Cook's-distance outliers
($d > 4/N$) removes 22 observations across 2 countries with negligible impact on
either elasticity. For the high-income subsample (17 EU countries above median GDP
per capita) only the EU elasticity ($\SampHiEUB$, p$=$\SampHiEUPval) is identified
and is not significant, consistent with minimal affordability constraints at higher
incomes. For the middle/lower-income subsample (16 countries including all EaP), EaP
elasticity ($\SampMidEaPB$, p$<$0.001) is somewhat smaller than the full-sample
estimate, reflecting within-group heterogeneity.

\input{tables/table8_sample_restrictions}

% ============================================================================
% REFERENCES
% ============================================================================

\bibliography{references}

\end{document}
