% ============================================================================
% Broadband Demand Elasticity Paper
% Journal: Elsevier Submission
% Author: Samir Orujov
% ============================================================================
%
% elsarticle document class for Elsevier journal submissions
% Options:
%   - preprint,review: double-spaced for review
%   - 12pt: font size
%   - authoryear: author-year citation style (requires natbib)
%   - 3p: three-column journal layout
%
% For final submission, use: \documentclass[final,3p,times,authoryear]{styles/elsarticle}
% For review submission, use current settings
%
\documentclass[preprint,review,12pt,authoryear]{styles/elsarticle}

% ============================================================================
% PACKAGES
% ============================================================================

% Line numbers for review (required for submission)
\usepackage{lineno}
\modulolinenumbers[5]

% Mathematics and symbols
\usepackage{amsmath,amssymb}
\usepackage{bbm}              % Indicator function \mathbbm{1}

% Tables and figures
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{float}
\usepackage{threeparttable}
\usepackage{array}
\usepackage{longtable}
\usepackage{adjustbox}       % For scaling tables to fit page width
\usepackage{pdflscape}       % For landscape tables if needed
\usepackage{setspace}        % For single spacing in tables

% Note: natbib is loaded automatically by elsarticle with authoryear option
% Set bibliography style
\bibliographystyle{styles/elsarticle-harv}

% Figure search path
\graphicspath{{figures/}}

% Hyperlinks (for electronic submission)
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}

% Journal information (update as needed for target journal)
\journal{Information Economics and Policy}

% Auto-generated macros (numbers from regression outputs)
% Regenerate with: python code/analysis/generate_paper_macros.py
\input{paper_macros}

% ============================================================================
% DOCUMENT
% ============================================================================

\begin{document}

% Suppress the default preprint footer ("Preprint submitted to ...") which can
% overlap with frontmatter in review mode. This is recommended when using
% review + line numbers.
\makeatletter
\def\ps@pprintTitle{\let\@oddhead\@empty\let\@evenhead\@empty
    \let\@oddfoot\@empty\let\@evenfoot\@empty}
\makeatother

\begin{frontmatter}

% Title
\title{The Transformation of Broadband Demand: From Discretionary Service to Essential Infrastructure (2010--2024)}

% Authors
\author[ada,icta,unec,ubs]{Samir Orujov\corref{cor1}}
\ead{sorujov@ada.edu.az}
\cortext[cor1]{Corresponding author}

\author[icta,charles]{Ilgar Ismayilov}

\author[icta]{Jeyhun Huseynzade}

% GitHub link moved to end note to reduce frontmatter height

\affiliation[ada]{organization={School of Business, ADA University},
                  city={Baku},
                  country={Azerbaijan}}

\affiliation[icta]{organization={Information and Communication Technology Agency},
                  city={Baku},
                  country={Azerbaijan}}

\affiliation[charles]{organization={Faculty of Social Sciences, Charles University},
                  city={Prague},
                  country={Czech Republic}}

\affiliation[unec]{organization={Azerbaijan State University of Economics (UNEC)},
                  city={Baku},
                  country={Azerbaijan}}

\affiliation[ubs]{organization={Université de Bretagne Sud},
                  city={Lorient},
                  country={France}}

% Abstract
\begin{abstract}
\input{sections/abstract}
\end{abstract}

% Keywords
\begin{keyword}
Broadband demand \sep Price elasticity \sep Panel data \sep Eastern Partnership \sep European Union \sep COVID-19 \sep Digital transformation
\end{keyword}

\begin{highlights}
\item Broadband price elasticity declined from $-0.61$ (EaP) and $-0.12$ (EU) to $\approx$0 by 2020
\item EaP countries were 5$\times$ more price-elastic than EU countries pre-COVID, then converged
\item Transformation began in 2015---not 2020---ruling out COVID-19 as the sole causal shock
\item GNI-based prices give significant results in all specs; PPP in only 25\% of EaP specs
\item Policy must shift from price subsidies to infrastructure deployment as demand matures
\end{highlights}

\end{frontmatter}

% Enable line numbers for the main text only (keeps frontmatter uncluttered)
\linenumbers

% ============================================================================
% MAIN SECTIONS
% ============================================================================

\section{Introduction}
\label{sec:introduction}
\input{sections/introduction}

\section{Literature Review and Theoretical Framework}
\label{sec:literature}
\input{sections/literature}

\section{Data and Sample}
\label{sec:data}
\input{sections/data}

\section{Empirical Methodology}
\label{sec:methodology}
\input{sections/methodology}

\section{Results}
\label{sec:results}
\input{sections/results}

\section{Discussion and Policy Implications}
\label{sec:discussion}
\input{sections/discussion}

\section{Conclusion}
\label{sec:conclusion}
\input{sections/conclusion}

% ============================================================================
% CREDIT AUTHOR STATEMENT (Elsevier requirement)
% ============================================================================

\section*{CRediT authorship contribution statement}
\textbf{Samir Orujov:} Conceptualization, Methodology, Software, Formal analysis, Data curation, Writing -- original draft, Writing -- review \& editing, Visualization.
\textbf{Ilgar Ismayilov:} Validation, Writing -- review \& editing, Supervision.
\textbf{Jeyhun Huseynzade:} Validation, Writing -- review \& editing.

% ============================================================================
% ACKNOWLEDGMENTS
% ============================================================================

\section*{Acknowledgments}
The authors thank Daniel Vertesy and Esperanza Magpantay of the ITU Statistics Division for clarifications on price basket definitions and data coverage. The authors also acknowledge the support of the Information and Communication Technologies Agency of the Republic of Azerbaijan. The views expressed in this paper are those of the authors and do not necessarily reflect the official positions of the affiliated institutions. Data were obtained from the International Telecommunication Union (ITU) and World Bank open data portals.

% ============================================================================
% DECLARATIONS
% ============================================================================

\section*{Declaration of competing interest}
The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

\section*{Data availability}
The data used in this study are publicly available from the International Telecommunication Union (ITU) World Telecommunication/ICT Indicators Database and the World Bank World Development Indicators. Replication code and processed analysis datasets are available at: \url{https://github.com/sorujov/Broadband-Demand-Elasticity}.

% ============================================================================
% APPENDICES
% ============================================================================

\clearpage
\appendix
\enlargethispage{4\baselineskip}

\section{Regression Tables}
\label{app:tables}
\setcounter{table}{0}
\small
\singlespacing
{\small\noindent This appendix presents the complete regression results referenced in the main text. All models use two-way fixed effects (country and year) with Driscoll--Kraay standard errors robust to heteroskedasticity, serial correlation, and cross-sectional dependence (bandwidth = 3).}
\vspace{-4pt}

\subsection{Baseline Pre-COVID Results}
\vspace{-2pt}
\small
\singlespacing

{ Table~\ref{tab:baseline} presents the baseline estimates for the pre-COVID period (2010--2019) across seven control specifications.}
\vspace{-2pt}

\input{tables/table1_baseline}
\clearpage

\subsection{COVID-19 Interaction Effects}
\small
\singlespacing

Table~\ref{tab:covid} presents results from the full sample (2010--2024) with COVID-19 period interactions.

\input{tables/table2_covid}
\clearpage

\subsection{Robustness to Alternative Price Measures}
% \small
\singlespacing

Table~\ref{tab:price_robustness} compares elasticity estimates across three price definitions: income-relative (baseline), PPP-adjusted, and nominal USD.

\input{tables/table3_price_robustness}
\clearpage

\subsection{Placebo Test for Pre-COVID Trends}
\small
\singlespacing

Table~\ref{tab:placebo} presents the placebo test splitting the pre-COVID period into early (2010--2014) and late (2015--2019) subperiods.

\input{tables/table4_placebo}
\clearpage

\section{Additional Robustness Checks}
\label{app:robustness}
\setcounter{table}{0}
\small
\singlespacing
\subsection{Period-Split: Acute vs.\ Post-Acute COVID}

Table~\ref{tab:period_split} decomposes the COVID period (2020--2024) into an acute phase
(2020--2021) and a post-acute phase (2022--2024). Both sub-periods exhibit near-zero
elasticity for EU and EaP countries, confirming that the disappearance of price sensitivity
is not driven solely by the acute pandemic shock. Panel~B excludes Ukraine and Belarus---whose
data reliability is compromised by war and sanctions---and reaches the same conclusion.

\input{tables/table5_period_split}
\clearpage

\subsection{IV Robustness: 2SLS Estimates}
\small
\singlespacing
Table~\ref{tab:iv_robustness} presents 2SLS estimates instrumenting fixed broadband price
with its own one-year lag ($\log$ fixed broadband price in GNI\%, lagged one year).

\textbf{Instrument rationale.}
National fixed broadband price baskets are set through annual regulatory and procurement
cycles, generating high year-to-year autocorrelation. The lagged price therefore strongly
predicts the current price (within-demeaned first-stage $F = 65$ pre-COVID; $F = 99$
full-sample), well above the Stock--Yogo 10\% maximal-IV-bias threshold of 16.38
\citep{stock2002testing}. The exclusion restriction is plausible: conditional on country
and year fixed effects plus time-varying controls, past prices affect current subscriptions
only through current prices, since broadband contracts are monthly or annual and consumers
cannot anticipate next year's regulated tariff when making adoption decisions today.\footnote{%
  We also tested international internet bandwidth as an alternative instrument. However,
  an over-identified specification combining both instruments strongly rejected the Hansen
  $J$ overidentification test ($\chi^2(2) = 262.89$, $p < 0.001$), indicating that
  bandwidth correlates with connection speed and quality, which directly enter the demand
  equation. Lagged price is therefore the sole preferred instrument.}

\textbf{Results.}
Panel~A of Table~\ref{tab:iv_robustness} shows pre-COVID results alongside the OLS
baseline. The 2SLS estimate is EU$=-0.172$ (SE$=$0.097, $p=$0.075) and EaP$=-0.615$
(SE$=$0.091, $p<$0.001)---essentially identical to OLS (EU$=-0.101$, EaP$=-0.595$),
confirming that simultaneity bias in the baseline TWFE estimates is negligible.
Panel~B uses the same instrument in the full-sample specification with COVID interactions:
EU$=-0.193$ (SE$=$0.058, $p<$0.001) and EaP$=-0.340$ (SE$=$0.088, $p<$0.001), consistent
with the COVID-era attenuation of price sensitivity found in OLS.

Overall, the IV results directly validate the OLS estimates: the endogeneity concern is
resolved and no evidence of upward simultaneity bias is found.

\input{tables/table6_iv_robustness}
\clearpage

\subsection{EaP Jackknife: Leave-One-Country-Out}
\small
\singlespacing
Table~\ref{tab:jackknife} reports leave-one-out estimates dropping each of the six EaP
countries in turn. EaP elasticity ranges from $-0.42$ to $-0.72$ across the six sub-samples.
Armenia is the most influential country: its exclusion shifts the EaP estimate from $-0.61$
to $-0.42$ (a change of 0.19 in magnitude), suggesting Armenia contributes relatively
low price-sensitivity to the aggregate. Excluding Belarus, which experienced political crisis
and sanctions during the sample period, yields a larger estimate ($-0.72$), consistent with
data reliability concerns for that country. Crucially, \textit{all} leave-one-out EaP
estimates are statistically significant at the 1\% level, and all confirm a substantial,
negative price elasticity. The EaP result is therefore not driven by any single country,
and the range $[-0.42, -0.72]$ reflects genuine heterogeneity within the EaP group
rather than measurement fragility.

\input{tables/table7_jackknife}
\clearpage

\subsection{Sample Restrictions}
\small
\singlespacing
Table~\ref{tab:sample_restrictions} reports robustness to alternative sample definitions.
The balanced-panel restriction (countries with complete 2010--2019 data) yields identical
estimates to the full sample, confirming that attrition is not driving the results.
Excluding outliers identified by Cook's distance ($d > 4/N$) removes 22 observations
across 2 countries with negligible impact on either elasticity estimate.
For the high-income subsample (17 EU countries above median GDP per capita), no EaP
countries are present, so only the EU elasticity ($\SampHiEUB$, p$=$\SampHiEUPval) is identified;
this estimate is not statistically significant, consistent with
higher-income consumers facing minimal affordability constraints.
For the middle/lower-income subsample (16 countries including all EaP), the EaP elasticity
($\SampMidEaPB$, p$<$0.001) is somewhat smaller than the full-sample estimate, reflecting
country heterogeneity at lower income levels.

\input{tables/table8_sample_restrictions}

% ============================================================================
% GENERATIVE AI DECLARATION
% ============================================================================

\section*{Generative AI and AI-assisted technologies in the writing process}
During the preparation of this work the authors used Claude (Anthropic) for programming and data-collection scripting, and for manuscript editing. After using this tool, the authors reviewed and edited all content as needed and take full responsibility for the content of this publication.

% ============================================================================
% REFERENCES
% ============================================================================

\bibliography{references}

\end{document}
