% ============================================================================
% Broadband Demand Elasticity Paper — ANONYMIZED MANUSCRIPT
%
% PURPOSE: Author-blinded version for double-blind peer review.
%          All author identifiers have been removed:
%            • Author names and affiliations omitted
%            • CRediT statement omitted
%            • Acknowledgments omitted (available upon acceptance)
%            • GitHub repository URL anonymised
%            • Institutional acknowledgment omitted
%          Upload this file as the "Manuscript file" in the submission system.
%          Upload titlepage.tex (separately compiled) as the "Title Page".
% ============================================================================

\documentclass[preprint,review,12pt,authoryear]{styles/elsarticle}

% ============================================================================
% PACKAGES
% ============================================================================

\usepackage{lineno}
\modulolinenumbers[5]

\usepackage{amsmath,amssymb}
\usepackage{bbm}

\usepackage{booktabs}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{float}
\usepackage{threeparttable}
\usepackage{array}
\usepackage{longtable}
\usepackage{adjustbox}
\usepackage{pdflscape}
\usepackage{setspace}

\bibliographystyle{styles/elsarticle-harv}
\graphicspath{{figures/}}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}

\journal{Economics of Innovation and New Technology}

\input{paper_macros}

% ============================================================================
% DOCUMENT
% ============================================================================

\begin{document}

\makeatletter
\def\ps@pprintTitle{\let\@oddhead\@empty\let\@evenhead\@empty
    \let\@oddfoot\@empty\let\@evenfoot\@empty}
\makeatother

\begin{frontmatter}

% ── Title ────────────────────────────────────────────────────────────────────
\title{Broadband Price Elasticity in EU and Eastern Partnership Countries:
       Evidence of a Decade-Long Transformation (2010--2024)}

% Authors and affiliations are omitted for double-blind review.
% See the separate Title Page file for full author information.

% ── Abstract ─────────────────────────────────────────────────────────────────
\begin{abstract}
\input{sections/abstract}
\end{abstract}

% ── Keywords ─────────────────────────────────────────────────────────────────
\begin{keyword}
Broadband demand \sep Price elasticity \sep Panel data \sep Eastern
Partnership \sep European Union \sep COVID-19 \sep Digital transformation
\end{keyword}

% ── Highlights ───────────────────────────────────────────────────────────────
\begin{highlights}
\item Broadband price elasticity declined from $\EaPBaseBr$ (EaP) and $\EUBaseBr$
      (EU) to $\approx$0 by 2020
\item EaP price elasticity $\EaPEURatio$$\times$ EU pre-COVID; demand-side
      harmonisation achieved by 2020
\item Transformation began in 2015---not 2020---ruling out COVID-19 as the sole
      causal shock
\item GNI-based prices give significant results in all specs; PPP in only
      \PctPPPSigEaP\% of EaP specs
\item Policy must shift from price subsidies to infrastructure deployment as
      demand matures
\end{highlights}

\end{frontmatter}

\linenumbers

% ============================================================================
% MAIN SECTIONS
% ============================================================================

\section{Introduction}
\label{sec:introduction}
\input{sections/introduction}

\section{Literature Review and Theoretical Framework}
\label{sec:literature}
\input{sections/literature}

\section{Data and Sample}
\label{sec:data}
\input{sections/data_anon}

\section{Empirical Methodology}
\label{sec:methodology}
\input{sections/methodology}

\section{Results}
\label{sec:results}
\input{sections/results}

\section{Discussion and Policy Implications}
\label{sec:discussion}
\input{sections/discussion}

\section{Conclusion}
\label{sec:conclusion}
\input{sections/conclusion}

% ============================================================================
% DECLARATIONS (anonymised for review)
% ============================================================================

% CRediT statement omitted — available upon acceptance.

% Acknowledgments omitted for double-blind review — available upon acceptance.

\section*{Funding}
This research did not receive any specific grant from funding agencies in the
public, commercial, or not-for-profit sectors.

\section*{Declaration of competing interest}
The authors declare that they have no known competing financial interests or
personal relationships that could have appeared to influence the work reported
in this paper.

\section*{Declaration of generative AI and AI-assisted technologies in the
manuscript preparation process}
During the preparation of this work the authors used Claude Sonnet 4.6
(Anthropic) for programming and data-collection scripting, and for manuscript
editing. After using this tool, the authors reviewed and edited all content as
needed and take full responsibility for the content of this publication.

\section*{Data availability}
The data used in this study are publicly available from the International
Telecommunication Union (ITU) World Telecommunication/ICT Indicators Database
and the World Bank World Development Indicators. Replication code and processed
analysis datasets are available at: [Repository link provided upon acceptance].

% ============================================================================
% REFERENCES
% ============================================================================

\bibliography{references}

% ============================================================================
% APPENDICES
% ============================================================================

\clearpage
\appendix
\enlargethispage{4\baselineskip}

\section{Regression Tables}
\label{app:tables}
\setcounter{table}{0}
\small
\singlespacing
{\small\noindent This appendix presents the complete regression results referenced
in the main text. All models use two-way fixed effects (country and year) with
Driscoll--Kraay standard errors robust to heteroskedasticity, serial correlation,
and cross-sectional dependence (bandwidth = 3).}
\vspace{-4pt}

\subsection{Baseline Pre-COVID Results}
\vspace{-2pt}
\small
\singlespacing

{Table~\ref{tab:baseline} presents the baseline estimates for the pre-COVID period
(2010--2019) across seven control specifications.}
\vspace{-2pt}

\input{tables/table1_baseline}
\clearpage

\subsection{COVID-19 Interaction Effects}
\small
\singlespacing

Table~\ref{tab:covid} presents results from the full sample (2010--2024) with
COVID-19 period interactions.

\input{tables/table2_covid}
\clearpage

\subsection{Robustness to Alternative Price Measures}
\singlespacing

Table~\ref{tab:price_robustness} compares elasticity estimates across three price
definitions: income-relative (baseline), PPP-adjusted, and nominal USD.

\input{tables/table3_price_robustness}
\clearpage

\subsection{Placebo Test for Pre-COVID Trends}
\small
\singlespacing

Table~\ref{tab:placebo} presents the placebo test splitting the pre-COVID period
into early (2010--2014) and late (2015--2019) subperiods.

\input{tables/table4_placebo}
\clearpage

\section{Additional Robustness Checks}
\label{app:robustness}
\setcounter{table}{0}
\small
\singlespacing

\subsection{Period-Split: Acute vs.\ Post-Acute COVID}

Table~\ref{tab:period_split} decomposes the COVID period (2020--2024) into an
acute phase (2020--2021) and a post-acute phase (2022--2024). Both sub-periods
exhibit near-zero elasticity for EU and EaP countries, confirming that the
disappearance of price sensitivity is not driven solely by the acute pandemic shock.
Panel~B excludes Ukraine and Belarus---whose data reliability is compromised by war
and sanctions---and reaches the same conclusion.

\input{tables/table5_period_split}
\clearpage

\subsection{IV Robustness: 2SLS Estimates}
\small
\singlespacing

Table~\ref{tab:iv_robustness} presents 2SLS estimates instrumenting fixed broadband
price with its own one-year lag ($\log$ fixed broadband price in GNI\%, lagged one year).

\textbf{Instrument rationale.}
National fixed broadband price baskets are set through annual regulatory and
procurement cycles, generating high year-to-year autocorrelation. The lagged price
therefore strongly predicts the current price (within-demeaned first-stage
$F = \IVFirstF$ pre-COVID; $F = \IVFullF$ full-sample), well above the Stock--Yogo
10\% maximal-IV-bias threshold of 16.38 \citep{stock2002testing}. The exclusion
restriction is plausible: conditional on country and year fixed effects plus
time-varying controls, past prices affect current subscriptions only through current
prices, since broadband contracts are monthly or annual and consumers cannot anticipate
next year's regulated tariff when making adoption decisions today.\footnote{We also
tested international internet bandwidth as an alternative instrument. However, an
over-identified specification combining both instruments strongly rejected the Hansen
$J$ overidentification test ($\chi^{2}(2) = 262.89$, $p < 0.001$), indicating that
bandwidth correlates with connection speed and quality, which directly enter the demand
equation. A further practical limitation is that ITU bandwidth data are missing for most
EU countries after 2018, and multiple imputation does not recover sufficient coverage to
support credible inference. Lagged price is therefore the sole preferred instrument.}

\input{tables/table6_iv_robustness}
\clearpage

\subsection{EaP Jackknife: Leave-One-Country-Out}

Table~\ref{tab:jackknife} reports leave-one-out estimates dropping each of the six
EaP countries in turn. EaP elasticity ranges from $-0.42$ to $-0.67$ across the six
sub-samples. Armenia is the most influential country: its exclusion raises the EaP
estimate from $-0.60$ to $-0.42$, suggesting Armenia contributes relatively high
price-sensitivity to the aggregate EaP estimate. Excluding Georgia yields the largest
negative estimate ($-0.67$), indicating that Georgia exhibits below-average price
sensitivity relative to the rest of the EaP group; removing it therefore shifts the
group-level estimate further from zero. Crucially, all leave-one-out EaP estimates are
statistically significant at the 1\% level, and all confirm a substantial, negative price
elasticity. The EaP result is therefore not driven by any single country, and the range
$[-0.42, -0.67]$ reflects genuine heterogeneity within the EaP group rather than
measurement fragility.

\input{tables/table7_jackknife}
\clearpage

\subsection{Sample Restrictions}

Table~\ref{tab:sample_restrictions} reports robustness to alternative sample
definitions. The balanced-panel restriction (countries with complete 2010--2019 data)
yields identical estimates to the full sample, confirming that attrition is not driving
the results. Excluding outliers identified by Cook's distance ($d > 4/N$) removes 22
observations across 2 countries with negligible impact on either elasticity estimate.
For the high-income subsample (17 EU countries above median GDP per capita), no EaP
countries are present, so only the EU elasticity ($-0.016$, $p=0.673$) is identified;
this estimate is not statistically significant, consistent with higher-income consumers
facing minimal affordability constraints. For the middle/lower-income subsample
(16 countries including all EaP), the EaP elasticity ($-0.433$, $p<0.001$) is somewhat
smaller than the full-sample estimate, reflecting country heterogeneity at lower income
levels.

\input{tables/table8_sample_restrictions}

\end{document}
