% ============================================================================
% TITLE PAGE
% Broadband Demand Elasticity Paper — Elsevier Submission
%
% PURPOSE: Standalone title page for double-blind peer review submission.
%          Upload this file as the "Title Page" item in the submission system.
%          The companion file paper_anonymous.tex is the author-blinded manuscript.
% ============================================================================

\documentclass[preprint,review,12pt,authoryear]{styles/elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}
\bibliographystyle{styles/elsarticle-harv}
\graphicspath{{figures/}}
\input{paper_macros}

\journal{Information Economics and Policy}

% Suppress default preprint footer
\makeatletter
\def\ps@pprintTitle{\let\@oddhead\@empty\let\@evenhead\@empty
    \let\@oddfoot\@empty\let\@evenfoot\@empty}
\makeatother

\begin{document}

\begin{frontmatter}

% ── Title ────────────────────────────────────────────────────────────────────
\title{Broadband Price Elasticity in EU and Eastern Partnership Countries:
       Evidence of a Decade-Long Transformation (2010--2024)}

% ── Authors ──────────────────────────────────────────────────────────────────
\author[ada,icta,unec,ubs]{Samir Orujov\corref{cor1}}
\ead{sorujov@ada.edu.az}
\cortext[cor1]{Corresponding author. Tel.: +99451 3171988}

\author[icta,charles]{Ilgar Ismayilov}

\author[icta]{Jeyhun Huseynzade}

% ── Affiliations ─────────────────────────────────────────────────────────────
\affiliation[ada]{organization={School of Business, ADA University},
                  city={Baku},
                  country={Azerbaijan}}

\affiliation[icta]{organization={Information and Communication Technology Agency},
                  city={Baku},
                  country={Azerbaijan}}

\affiliation[charles]{organization={Faculty of Social Sciences, Charles University},
                  city={Prague},
                  country={Czech Republic}}

\affiliation[unec]{organization={Azerbaijan State University of Economics (UNEC)},
                  city={Baku},
                  country={Azerbaijan}}

\affiliation[ubs]{organization={Universit\'{e} de Bretagne Sud},
                  city={Vannes},
                  country={France}}

% ── Abstract ─────────────────────────────────────────────────────────────────
\begin{abstract}
\input{sections/abstract}
\end{abstract}

% ── Keywords ─────────────────────────────────────────────────────────────────
\begin{keyword}
Broadband demand \sep Price elasticity \sep Panel data \sep Eastern
Partnership \sep European Union \sep COVID-19 \sep Digital transformation
\end{keyword}

\end{frontmatter}

% ── CRediT Author Contribution Statement ─────────────────────────────────────
\section*{CRediT authorship contribution statement}
\textbf{Samir Orujov:} Conceptualization, Methodology, Software, Formal
analysis, Data curation, Writing -- original draft, Writing -- review \&
editing, Visualization.
\textbf{Ilgar Ismayilov:} Validation, Writing -- review \& editing, Supervision.
\textbf{Jeyhun Huseynzade:} Validation, Writing -- review \& editing.

% ── Acknowledgments ──────────────────────────────────────────────────────────
\section*{Acknowledgments}
The authors thank Daniel Vertesy and Esperanza Magpantay of the ITU Statistics
Division for clarifications on price basket definitions and data coverage. The
authors also acknowledge the support of the Information Communication
Technologies Agency of the Republic of Azerbaijan. The views expressed in this
paper are those of the authors and do not necessarily reflect the official
positions of the affiliated institutions. Data were obtained from the International
Telecommunication Union (ITU) and World Bank open data portals.

% ── Declarations ─────────────────────────────────────────────────────────────
\section*{Funding}
This research did not receive any specific grant from funding agencies in the
public, commercial, or not-for-profit sectors.

\section*{Declaration of competing interest}
The authors declare that they have no known competing financial interests or
personal relationships that could have appeared to influence the work reported
in this paper.

\section*{Declaration of generative AI and AI-assisted technologies in the
manuscript preparation process}
During the preparation of this work the authors used Claude (Anthropic) for
programming and data-collection scripting, and for manuscript editing. After
using this tool, the authors reviewed and edited all content as needed and take
full responsibility for the content of this publication.

\section*{Data availability}
The data used in this study are publicly available from the International
Telecommunication Union (ITU) World Telecommunication/ICT Indicators Database
and the World Bank World Development Indicators. Replication code and processed
analysis datasets are available at:
\url{https://github.com/sorujov/Broadband-Demand-Elasticity}.

\end{document}
