This paper documents a fundamental transformation in broadband demand across 33 European countries over 2010--2024. Three findings emerge from our two-way fixed effects analysis with Driscoll--Kraay standard errors.

First, substantial regional heterogeneity characterized pre-COVID demand. Eastern Partnership countries exhibited strong price sensitivity ($\varepsilon = \EaPBaseBr$, p$<$0.001)---five times larger than EU elasticity ($\varepsilon = \EUBaseBr$, p$<$0.05)---reflecting income-driven affordability constraints. Results are robust across \RobNSpecs{} specifications.

Second, both regions converged to near-zero elasticity during 2020--2024, with price changes having no detectable effect on adoption. This shift indicates broadband's transformation from discretionary service to essential necessity.

Third, and most critically, this transformation predates COVID-19. EU year-by-year estimates reveal elasticity declining from $\EUYbyYRefElast$ in 2015 toward zero, and placebo tests confirm EaP elasticity was also declining before COVID-19 (p$=$\PlaceboEaPPval)---a result that requires the pooled placebo design given EaP's limited sample of five to six countries. The pandemic accelerated but did not initiate broadband's integration into economic and social life.

These findings yield four contributions: updated elasticity estimates for 2010--2024; evidence of regional heterogeneity within Europe; demonstration that demand elasticity is time-varying; and reframing of COVID-19's impact as acceleration rather than shock. Methodologically, we show price measurement critically affects inference---income-relative prices yield robust results while PPP-adjusted prices do not.

For policy, the implications are direct. When demand was elastic (pre-2015), affordability interventions---subsidies, price regulation---effectively expanded adoption. As elasticity approached zero, these tools became ineffective. Policy must now prioritize infrastructure deployment and universal service obligations over price-based mechanisms. The near-zero elasticity suggests broadband resembles utilities like water or electricity, potentially justifying stronger regulatory oversight and public investment to ensure equitable access.

Future research should examine household-level heterogeneity to identify remaining price-sensitive populations, extend analysis to mobile broadband and other regions, and develop quality-adjusted price indices. As broadband completes its transformation into essential infrastructure, understanding these dynamics becomes crucial for bridging digital divides worldwide.
