\subsection{Data Sources}

We construct a balanced panel of 33 countries observed annually from 2010 through 2024, combining data from the International Telecommunication Union (ITU) and World Bank. Our sample includes 27 European Union member states and 6 Eastern Partnership countries (Armenia, Azerbaijan, Belarus, Georgia, Moldova, and Ukraine), totaling 495 country-year observations.

\textbf{Telecommunications variables} come from ITU World Telecommunication/ICT Indicators Database \citep{itu2024data}. The dependent variable is fixed broadband subscriptions per 100 inhabitants, measuring adoption intensity. Price variables include three measures: (1) price of the ITU fixed broadband basket as percentage of monthly GNI per capita (income-relative price), (2) price in PPP-adjusted US dollars, and (3) price in nominal US dollars. Additional ICT variables include international internet bandwidth (Gbit/s) and mobile cellular subscriptions per 100 inhabitants.

\textbf{Economic and institutional variables} come from World Bank World Development Indicators and Worldwide Governance Indicators \citep{worldbank2024wdi, worldbank2024wgi}. Key controls include GDP per capita (current US\$), GDP growth rate, inflation, urban population percentage, tertiary education enrollment, regulatory quality index, R\&D expenditure as \% of GDP, and secure internet servers per million people. These variables capture economic development, human capital, institutional quality, and digital infrastructure.

\subsection{Sample Construction and Balance}

Our sample selection prioritizes internal validity over external generalizability. We restrict to countries with complete data for all years 2010--2024, ensuring a balanced panel without compositional changes that could confound temporal comparisons. This is particularly important given our focus on time-varying elasticity: different countries entering and exiting the sample would generate spurious trends.

The 33-country sample represents substantial variation in both broadband adoption and economic development. Fixed broadband penetration ranges from 3.2 subscriptions per 100 (Armenia, 2010) to 48.9 (France, 2024). GDP per capita ranges from \$2,094 (Ukraine, 2015) to \$137,517 (Luxembourg, 2024). Broadband prices show similar heterogeneity: income-relative prices (as \% of GNI per capita) range from 0.31\% (Austria, 2023) to 11.52\% (Armenia, 2010).

Table~\ref{tab:descriptives} presents descriptive statistics for the full sample and separately for EU and EaP regions. Several patterns emerge. First, EaP countries have substantially lower broadband penetration (mean \DescSubsEaP{} vs.\ \DescSubsEU{} subscriptions per 100) and GDP per capita (mean \$\DescGDPEaP{} vs.\ \$\DescGDPEU). Second, income-relative prices are higher in EaP (mean \DescPriceGNIEaP\% of GNI vs.\ \DescPriceGNIEU\%), reflecting lower incomes despite similar absolute prices. Third, EaP countries show higher variance in key variables, consistent with greater development heterogeneity.

Importantly, both regions exhibit substantial within-country variation over time---the source of identifying variation for our fixed effects models. Income-relative prices declined 45\% on average from 2010 to 2024, with within-country standard deviation of 0.5 percentage points (pre-COVID) and 0.3 percentage points (COVID period). This variation is essential for identifying price elasticity in models with country and year fixed effects.

\subsection{Variable Construction}

Following standard practice in telecommunications demand studies \citep{grzybowski2015fixed, madden2015demand}, we log-transform all continuous variables to estimate constant-elasticity specifications:
\begin{equation}
    \ln(\text{subscriptions}_{it}) = \beta \ln(\text{price}_{it}) + \mathbf{X}_{it}'\gamma + \alpha_i + \delta_t + \varepsilon_{it}
\end{equation}
where the coefficient $\beta$ directly represents price elasticity. This transformation is particularly appropriate for broadband demand given theoretical predictions of constant elasticity \citep{hausman2001private}.

We create several derived variables for robustness checks. We construct regional interaction terms allowing heterogeneous price effects for EU versus EaP countries. For COVID-19 analysis, we create period dummies (pre-COVID: 2010--2019; COVID: 2020--2024) and interaction terms testing for structural breaks.

\subsection{Data Quality and Limitations}

Several data quality considerations warrant discussion. First, ITU price data represent
standard baskets that enable consistent cross-country comparisons \citep{itu2024data}.
An important measurement consistency issue arises from the ITU's revision of its basket
definition: the minimum data threshold was raised from 1\,GB to 5\,GB during the sample
period. Our price series combines both successive basket definitions: the 1\,GB basket (indicator I154\_FBB) for 2010--2017 and the 5\,GB basket (indicator I154\_FBB5) for 2018--2024, matching the ITU's own transition year. The basket change reflects that a 5\,GB monthly allocation had become the prevailing entry-level offering by 2018 rather than a shift in observed price levels. Consistent with this interpretation, the series shows no discontinuity at the handoff: the cross-country median income-relative price was 1.22\% of GNI in 2017 and 1.27\% in 2018, well within normal year-to-year variation. Data coverage and
basket definitions were verified directly with Daniel Vertesy and Esperanza Magpantay of
the ITU Statistics Division (personal communication, December 2025). Standardised baskets may not
capture marginal prices for high-consumption or bundled plans, but they represent the
policy-relevant entry-level price that is most responsive to affordability constraints.

Second, some variables exhibit missing values, particularly for earlier years.
Before imputation, the dependent variable (fixed broadband subscriptions) and the
primary price measure (\% of GNI per capita) have fewer than 0.5\% missing observations.
The most affected control variables are R\&D expenditure (12.7\% missing) and tertiary
education enrollment (9.5\% missing), with missingness concentrated in the early years
2010--2013 for smaller EaP countries. We address missingness through within-country linear interpolation
for interior gaps and forward-filling for leading/trailing values, following
standard panel data practice. All imputed values are flagged internally, and the
main elasticity estimates are unchanged when the analysis is restricted to non-imputed
observations only (EU: $-0.11$; EaP: $-0.60$; differences from Table~\ref{tab:baseline}
are below 0.02 in magnitude).

Third, the COVID-19 period presents measurement challenges as usage patterns shifted dramatically \citep{oecd2021covid}. We address this by examining within-country price variation rather than relying on cross-country comparisons, which could be confounded by differential pandemic severity or policy responses. Additionally, our placebo test directly examines whether observed COVID-period changes reflect continuation of pre-existing trends.

Fourth, Eastern Partnership countries experienced various shocks during the sample period including political transitions, conflicts, and currency crises. While country fixed effects absorb time-invariant country characteristics, time-varying shocks could confound estimates. We address this through year fixed effects (absorbing common shocks) and robustness checks examining alternative subperiods (see Section~\ref{sec:results}).

Fifth, and most importantly for the COVID-period analysis, two EaP countries face severe data reliability concerns in 2022--2024. \textbf{Ukraine} has experienced large-scale infrastructure destruction following the February 2022 invasion, which plausibly distorts ITU-reported broadband subscription and price data for that period. \textbf{Belarus} has been subject to broad international sanctions and deep political crisis since 2020, which may affect both the quality of ITU data reporting and the normal functioning of broadband markets. Since Ukraine and Belarus constitute two of the six EaP countries (33\% of the group), their inclusion in the 2020--2024 period could bias the EaP COVID-period elasticity estimates. We address this concern in Section~\ref{sec:results} by reporting robustness checks that (a) exclude Ukraine and Belarus entirely from COVID-period regressions and (b) drop their 2022--2024 observations while retaining 2020--2021. The key finding of near-zero EaP elasticity during the COVID period is robust to both exclusions.

\input{tables/table_descriptives}
